---
title: "Getting started with ambolt"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ambolt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## What is ambolt?

ambolt is for R developers who want to build web apps without learning a
frontend framework. You write R functions that describe your layout, your
inputs, and your outputs. The framework handles the frontend — generating a
modern single-page application with interactive components, responsive
layout, and URL routing.

If you've used Shiny, the workflow will feel familiar: you define **what**
exists (inputs, outputs) and **where** it goes (layout), then start the app.
The key difference is that ambolt compiles a real frontend at startup rather
than rendering HTML on every interaction — which makes apps faster and
easier to extend.

## Prerequisites

- **R** (>= 4.1) with the ambolt package installed
- **Node.js** (>= 18) — ambolt uses npm behind the scenes to build the
  frontend on first startup

Install ambolt from GitHub:

```{r}
# install.packages("remotes")
remotes::install_github("LCHansson/ambolt")
```

## Your first app: Hello World

Create a file called `app.R`:

```{r}
library(ambolt)

app <- create_app()

# A text input in the sidebar
app$input("name", "text", label = "Your name", placeholder = "Type here...")

# An HTML output in the main area
app$output("greeting", "html",
  depends_on = "name",
  render = function(params) {
    sprintf("<h2>Hello, %s!</h2>", params$name)
  }
)

# Place the input in the sidebar, the output in the main area
app$ui(sidebar_layout(
  sidebar("name"),
  main("greeting")
))

app$run()
```

Run it:

```bash
Rscript app.R
```

Visit <http://localhost:3000>. You'll see a sidebar with your text input on
the left and the greeting on the right — updating as you type. The first
startup takes a few seconds (npm installs dependencies and compiles the
frontend); subsequent starts use a cache and are near-instant.

Let's walk through what happened:

1. `create_app()` creates an app object — a container for everything.
2. `app$input()` declares a text input with an id, type, and label.
3. `app$output()` declares an output that depends on the input and renders
   HTML from the current input values.
4. `app$ui()` with `sidebar_layout(sidebar(...), main(...))` tells the
   framework where to place things — inputs on the left, outputs on the right.
5. `app$run()` starts the server and opens the browser.

## Adding a plot

Extend the app with a scatter plot that responds to user inputs:

```{r}
library(ambolt)

app <- create_app()

app$input("n", "slider", label = "Number of points",
  value = 50, min = 10, max = 500, step = 10)

app$input("color", "select", label = "Color",
  choices = c("steelblue", "coral", "forestgreen"))

app$output("scatter", "plot",
  depends_on = c("n", "color"),
  render = function(params) {
    n <- as.numeric(params$n)
    ggplot2::ggplot(
      data.frame(x = rnorm(n), y = rnorm(n)),
      ggplot2::aes(x, y)
    ) + ggplot2::geom_point(color = params$color, alpha = 0.6)
  }
)

app$ui(sidebar_layout(
  sidebar("n", "color"),
  main("scatter")
))

app$run()
```

The plot output renders as inline SVG — no image files, no file I/O. The
framework calls your render function whenever the depended-upon inputs
change.

## Layout building blocks

The layout functions are composable R functions that nest however you want.
Here are the most common ones:

**`section()`** groups related items with an optional label:

```{r}
sidebar(
  section("Vehicle", "make", "model", "year"),
  section("Economics", "fuel_price", "discount_rate")
)
```

**`columns()`** puts children side by side (collapses to stacked on mobile):

```{r}
main(
  columns("chart_a", "chart_b", gap = "2rem")
)
```

**`details()`** creates a collapsible section:

```{r}
main(
  "result",
  details("Show assumptions",
    "assumption_1", "assumption_2", "assumption_3"
  )
)
```

These are plain R functions that return lists. You can nest them, store them
in variables, pass them to other functions — whatever makes your layout code
readable.

## Dev mode

During development, set the `AMBOLT_DEV` environment variable to get a live-
reloading development server:

```bash
AMBOLT_DEV=TRUE Rscript app.R
```

| | Dev mode | Production mode |
|---|---|---|
| **Frontend** | Vite dev server with hot reload | Pre-built static files |
| **Speed** | Rebuilds on file changes | Cached, near-instant startup |
| **Use for** | Development | Deployment |

## Next steps

Your app has one page. For multi-page apps with navigation, modals, and
authentication, see `vignette("multi-page-with-auth", package = "ambolt")`.

For a real-world example, see
[Elektrifieringskollen](https://elektrifieringskollen.se) — a production
fleet cost calculator built with ambolt.
