---
title: "Building a multi-page app with authentication"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a multi-page app with authentication}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## When one page isn't enough

The getting-started vignette covers single-page apps with a sidebar and
main area. But real applications grow: you need separate screens, a
navigation menu, login pages, pop-up forms, and backend structure that
scales.

ambolt's multi-page mode gives you all of this. This vignette walks
through the patterns used in a production six-page CRM built with ambolt.

## Pages and navigation

Instead of `app$input()` / `app$output()` / `app$ui()`, multi-page apps
use `app$page()` and `app$module()`:

```{r}
library(ambolt)

app <- create_app()
app$meta(title = "My App", lang = "en")

app$page("overview", title = "Overview", icon = "house",
  ui = page_content(
    page_header("Overview", subtitle = "At a glance"),
    stat_cards("overview/stats")
  )
)

app$page("contacts", title = "Contacts", icon = "people",
  ui = page_content(
    page_header("Contacts", actions = list(
      create_button("New contact", modal = "contacts/create", icon = "plus")
    )),
    "contacts/table"
  )
)

app$run()
```

Each page gets an entry in the sidebar navigation — with an icon, a label,
and URL hash routing (`#/overview`, `#/contacts`). The `ui` argument takes a
layout tree built from the same composable functions as single-page apps:
`page_content()`, `page_header()`, `section()`, `columns()`, and so on.

Notice that some children are strings like `"overview/stats"` and
`"contacts/table"`. These reference **module outputs** — components that
fetch their own data from the backend. We'll define those next.

## Modules: organize your backend

A module groups related endpoints, outputs, and modals under a namespace:

```{r}
app$module("overview", function(mod, app) {
  # Endpoint: GET /api/overview/stats
  mod$get("/stats", function(req, res) {
    n <- DBI::dbGetQuery(app$context$db, "SELECT COUNT(*) AS n FROM contacts")$n
    res$json(list(total = n, active = n))
  })

  # Declare the output so the page can reference it as "overview/stats"
  mod$output("stats", type = "stats",
    endpoint = "/api/overview/stats",
    cards = list(
      list(key = "total", label = "Total", color = "primary"),
      list(key = "active", label = "Active", color = "success")
    )
  )
})
```

The module function receives a `mod` object with the same HTTP verb methods
as the app (`mod$get()`, `mod$post()`, etc.), but paths are automatically
prefixed with `/api/{module_id}`. This keeps your backend organized as the
app grows — each page gets its own module with its own endpoints.

## Authentication

ambolt has a built-in authentication system. You provide two functions —
one to verify a session token, one to handle login — and the framework
takes care of the login page, cookies, and route protection:

```{r}
app$auth(
  verify = function(token, db) {
    # Look up the session token in your database
    # Return user data (list) or NULL if invalid
    user <- get_session_user(db, token)
    if (is.null(user)) return(NULL)
    list(id = user$id, name = user$name, is_admin = user$is_admin)
  },

  login = function(username, password, db) {
    # Verify credentials against your database
    user <- verify_user(db, username, password)
    if (is.null(user)) return(NULL)
    # Create a session and return token + user data
    token <- ambolt_session_token()
    create_session(db, user$id, token)
    list(token = token, user = list(id = user$id, name = user$name))
  }
)
```

When auth is configured:

- Every `/api/` route is protected — unauthenticated requests get a 401
- The framework serves a login page at the app root
- Session cookies are HttpOnly and SameSite=Strict
- A rate limiter blocks brute-force login attempts
- The `req$user` object is available in every route handler

The auth system is pluggable: you bring your own database and verify
function. ambolt provides helpers for the common operations:

- `ambolt_session_token()` — generates a cryptographically random token
- `ambolt_hash_password(password)` — hashes a password (SHA-256)
- `ambolt_verify_password(password, hash)` — verifies a password against
  a stored hash

## Modals

Modals are pop-up dialogs for forms, detail views, and confirmations.
They can use the same layout functions as pages:

```{r}
app$module("contacts", function(mod, app) {
  # Form modal
  mod$modal("create", size = "lg", render = function(params) {
    list(
      title = "New contact",
      fields = list(
        list(id = "name", type = "text", label = "Name", required = TRUE),
        list(id = "email", type = "text", label = "Email")
      ),
      submit = list(
        endpoint = "/api/contacts",
        method = "POST",
        emit = "contacts:updated",
        label = "Create"
      )
    )
  })

  # Detail modal with composable layout
  mod$modal("profile", size = "lg", render = function(params) {
    contact <- get_contact(app$context$db, params$id)
    list(
      title = contact$name,
      content = page_content(
        page_header(contact$name, subtitle = contact$role),
        section("Details",
          html_block(sprintf("<p><strong>Email:</strong> %s</p>", contact$email))
        )
      )
    )
  })
})
```

There are three modal shapes:

- **Form modals** — return `list(title, fields, submit)`. The framework
  renders the form inputs and handles submission automatically.
- **Content modals** — return `list(title, content)` where `content` is
  a layout tree. The framework renders it using the same components as
  pages.
- **HTML modals** — return `list(title, html)` for one-off custom content.

Content modals are the most powerful: they render at request time, so the
layout can depend on database queries, parameters, and user context. The
`html_block()` function serves as an escape hatch for anything the built-in
layout functions don't cover.

## Theming

Set your brand colors once, and every component picks them up:

```{r}
app$theme(
  colors = list(
    primary = "#006589",
    secondary = "#710049",
    success = "#339933",
    warning = "#E37222"
  )
)
```

The theming system works in three tiers:

1. **Theme tokens** (global) — `app$theme()` sets CSS variables that
   propagate to all components.
2. **Semantic props** (per-component) — arguments like `compact`, `gap`,
   `variant` that control layout behavior.
3. **Escape hatch** (per-instance) — `class` and `style` arguments on any
   layout function for one-off overrides.

For full control, use `app$theme(css = "...")` to inject arbitrary CSS.

## Putting it together

Here's the skeleton of a complete multi-page app with auth and theming:

```{r}
library(ambolt)

app <- create_app()
app$meta(title = "My CRM", lang = "en")
app$theme(colors = list(primary = "#006589"))

# Database connection (stored in app$context for access in handlers)
app$context$db <- DBI::dbConnect(RSQLite::SQLite(), "my_app.sqlite")
app$on_stop(function() DBI::dbDisconnect(app$context$db))

# Authentication
app$auth(
  verify = function(token, db) { ... },
  login = function(username, password, db) { ... }
)

# Pages
app$page("overview", title = "Overview", icon = "house",
  ui = page_content(
    page_header("Overview"),
    stat_cards("overview/stats")
  )
)

app$page("contacts", title = "Contacts", icon = "people",
  ui = page_content(
    page_header("Contacts"),
    "contacts/table"
  )
)

# Modules (endpoints, outputs, modals)
app$module("overview", function(mod, app) { ... })
app$module("contacts", function(mod, app) { ... })

app$run()
```

This is the same pattern used in production apps built with ambolt —
multi-page CRMs with form modals, database migration frameworks, and
export pipelines.

## Under the hood

*You don't need to know any of this to use ambolt.* But if you're curious,
or want to extend the framework:

ambolt generates a [Svelte 5](https://svelte.dev/) frontend from your R
declarations. For pages, the R codegen walks your layout tree at build time
and emits static Svelte markup, which is compiled by Vite into optimized
JavaScript. For modals, the layout tree is serialized as JSON and walked
at request time by a recursive Svelte component called `RenderNode`.

The 30+ Svelte components live in `inst/svelte/components/`. You can add
new component types by defining an R constructor (with `.drop_nulls()`),
exporting it, and adding a dispatch branch in `RenderNode.svelte`.
